---
title: "Financial Time series Analysis and Forecasting using R"
author: "Daniel James" 
date: "`r Sys.Date()`"
#output: html_document
output: github_document
---
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "",
                      collapse = TRUE,
                      echo = TRUE,
                      error = TRUE, # do not interrupt in case of errors
                      message = FALSE,
                      warning = FALSE,
                      comma <- function(x) format(x, digits = 2, big.mark = ",")
)
````
# Stock Forecasting with Google
### Outline
###### Time series is a sequential set of data points, measured typically over successive times. It is mathematically defined as a set of vectors {Yt,t∈Z} where t represents the time. The data used for all the below method (ARIMA) were historical daily prices of the Google. The goal of this study is to model and forecast financial time series using the stochastic process (ARIMA).

### Load Packages
#### Important librarys
##### The packages used in this study are listed below.

### Step 1 - Install necessary packages

```{r install_packages, include = TRUE}
library(pacman)


p_load(lubridate)                 # for dates manipulation 
p_load(fBasics)                   # to enhanced summary statistics
p_load(lmtest)                    # for coefficients significance tests
p_load(urca)                      # for unit rooit test
p_load(ggplot2)                   # for data visualization
p_load(quantmod)                  # for getting financial data
p_load(PerformanceAnalytics)      # to calculating returns
p_load(rugarch)                   # for GARCH modeling
p_load(FinTS)                     # for ARCH test
p_load(forecast)                  # for ARMA modeling
p_load(strucchange)               # for structural changes 
p_load(TSA)                       # for ARMA order identification 
p_load(tseries)                   # for others
p_load(timeSeries)                # for others
p_load(timeSeries)                # for others
p_load(xts)                       # for others
p_load(pastecs)                   # for others
p_load(readxl)                    # to load in data
p_load(zoo)                       # for others
````
### Step 2   Import Data from Yahoo! Finance
#### Getting Data
###### The [Google Stock dataset](https://online.stat.psu.edu/onlinecourses/sites/stat501/files/ch15/google_stock.txt) having a total number of 105 observations and 7 variables. The data downloaded from [Yahoo! Finance](https://finance.yahoo.com/?guccounter=1&guce_referrer=aHR0cHM6Ly9yc3R1ZGlvLXB1YnMtc3RhdGljLnMzLmFtYXpvbmF3cy5jb20v&guce_referrer_sig=AQAAAJH4MynrY4bQqPd-bmWidSVVMB6f_jCjAijLbw8QkktShpT9tDXal9DA6hemxvOm04KVLWBfAwdlWugkljL09B5LCOZSzKPBbO5KuJeN6fFUB9LqHhrIArqBn_0MT8wygguV8ZDJ2m4K7waoyOoSVuRjqOS2uyjLfJ9ttWOUo2Rk). The closing price was chosen to be modeled and predicted. Which are covered the period from February 7, 2005 to July 7, 2005. We will analyze the dataset to identify the order of an autoregressive model.

###### Taking advantage of the getSymbols() function made available within quantmod package we get Google stock.
### Step 2 - Import Data

```{r Import_Google_data_from_cloud, include = TRUE}
GOOGLE = getSymbols('GOOG', from='2005-02-07', to='2005-07-08',auto.assign = FALSE)
````

```{r omit_na, include = TRUE}
GOOGLE = na.omit(GOOGLE)
head(GOOGLE)
````

#### Select 


```{r select_variable, include = TRUE}
# Select the relevant close price series
stock_prices = GOOGLE[,4]
head(stock_prices)
````

```{r print_tail, include = TRUE}
tail(stock_prices)
````

### Import data from laptop

###### The Google Stock dataset downloded from: [https://online.stat.psu.edu](https://online.stat.psu.edu)

```{r import_data_from_laptop, include = TRUE}
google_stock <- read_excel("google_stock.xlsx")
# first 6 observations
head(google_stock)
````

```{r print_tail_from_imported_data, include = TRUE}
tail(google_stock)
````

### Step 3 - Descriptive statistics

```{r print_descriptive_stat, include = TRUE}
summary(google_stock)
````

###### For more descriptive statistics, use stat.desc() from the package {pastecs}:

```{r print_descriptive_stat_from_pastecs_pkg, include = TRUE}
stat.desc(google_stock)
````
### Visualize the time series

###### More precisely, we have available OHLC (Open, High, Low, Close) index value, adjusted close value and trade volume. Here we can see the corresponding chart as produced by the chartSeries within the quantmod package.

```{r visualize_data_in_style, include = TRUE}
chartSeries(GOOGLE, type = "bars", theme="white",main="Google Stock")
````



